---
title: "PS270 Final Project"
author: "Margaret Tate"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project Thoughts 

I am interested in exploring data related to... trends in populism or democratic backsliding globally.  

## Project Milestone 2  
My research question for my final project is interested in populism, specifically: is populist rhetoric more prominent in leaders campaign speeches then other kinds of speeches? I hypothesize that if leaders are making campaign speeches, then they use more populist rhetoric, and use it more often, then in other types of speeches. The explanatory variable in this project is speech type, which in the data set as a whole can be classified as campaign, ribbon cutting, international, or famous. This is a categorical variable, determined by classifying the context of speeches a leader makes. The outcome variable is populist rhetoric which is this case is measured by a populism score determined by one or two coders using "holistic grading," according to Harvard Dataverse. While looking through certain leaders in the data, I noticed that campaign speeches appeared to have high populism scores, and wanted to see if that was a universally applicable trend. One observed trend that could disprove my hypothesis that populism scores seemed much more correlated by leader then by type of speech in some sections I was looking at. 

```{r}
populism <- read.csv("FinalProjectPopulismData2.csv")
```

```{r}
library(tidyverse)
```

```{r}
populism$totalaverage[populism$totalaverage == "NA"] <- NA
populism$speechtype <- tolower(populism$speechtype)
populism$speechtype <- trimws(populism$speechtype) 
populism$speechtype[populism$speechtype %in% c("ribboncutting", "ribbon cutting")] <- "ribbon cutting"##needed to use AI to fix some issues with discrepancies in data and it made these suggetions
populism$totalaverage <- as.numeric(populism$totalaverage)
populism1 <- populism |>
  group_by(speechtype) |>
  summarize(mean_totalaverage = mean(totalaverage, na.rm = TRUE))
populism_by_speechtype <- ggplot(data = populism1, aes(x = speechtype, y = mean_totalaverage)) +
  geom_col(fill = "dodgerblue4") +
  labs(title = "Mean Populism Score by Speech Type",
       x = "Speech Type",
       y = "Mean Score")
populism_by_speechtype
```

```{r}
populism1
```

```{r}
populism$totalaverage[populism$totalaverage == "NA"] <- NA
populism$speechtype <- tolower(populism$speechtype)
populism$speechtype <- trimws(populism$speechtype) 
populism$speechtype[populism$speechtype %in% c("ribboncutting", "ribbon cutting")] <- "ribbon cutting"##needed to use AI to fix some issues with discrepancies in data and it made these suggetions
populism$totalaverage <- as.numeric(populism$totalaverage)
populism1 <- populism |>
  group_by(speechtype,region) |>
  summarize(mean_totalaverage = mean(totalaverage, na.rm = TRUE))
populism_by_speechtype <- ggplot(data = populism1, aes(x = speechtype, y = mean_totalaverage)) +
  geom_col(fill = "dodgerblue4") +
  facet_wrap(~region) +
  labs(title = "Mean Populism Score by Speech Type and Region",
       x = "Speech Type",
       y = "Mean Score")
populism_by_speechtype
```

Looking at the first figure, it seems like I need to change the research question I am asking. Instead I want to find out if populist rhetoric has become more prominent over time? I hypothesize that populist rhetoric has become more prominent over time after observing global democratization over time. 

```{r}
populism$totalaverage <- as.numeric(populism$totalaverage)
populism2 <- populism |>
  group_by(yearbegin) |> 
  filter(yearbegin >1995) |>
  summarize(mean_totalaverage = mean(totalaverage, na.rm = T))
populism_by_year <- ggplot(data = populism2, aes(x = yearbegin, y = mean_totalaverage)) + 
    geom_line(color = "dodgerblue4") +
  geom_smooth(se = FALSE, color = "red", linewidth = 0.5) +
  labs(title = "Mean Populism Score by Year", x = "Year Candidates Term Began", y = "Mean Populism Score" )
  
populism_by_year
```

This shows that there has been a clear upward trend between 1995 and 2020 in the populist rhetoric recorded in this data set, which a relatively stagnant period between 2005 and 2015 and a sharp acceleration upwards after 2015, suggesting increasing prevalence in the last decade. 

